---
title: "Project1_olgol-19"
author: "olgol-19"
date: "October 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

Data was provided by **Urban Ministries of Durham**.

Urban Ministries of Durham (UMD) operates the Food Pantry and Clothing Closet and Community Cafe to address basic emergency needs for food, clothing, diapers and hygiene supplies.

The quistions to be answered:

1. The total number of families supported?
2. The number of families supported by year?
3. The food pounds provided by year?
4. The number of diapers provided by year?
5. The number of clothing items provided by year?
6. The number of school kits and hygien kits provided by year?
7. Average family size (by column food provided for)? 


# Data cleaning
I will use only non-character columns for analysis.
```{r message=FALSE, warning=FALSE}

UMD_data <- read_tsv("https://raw.githubusercontent.com/biodatascience/datasci611/gh-pages/data/project1_2019/UMD_Services_Provided_20190719.tsv") %>%
  mutate(N_Date = as.Date(Date, "%m/%d/%Y")) %>%
  select_if(~!is.character(.)) %>%
  select_if(~!is.logical(.)) %>%
  arrange(N_Date)

ggplot(data = UMD_data, aes(x=N_Date)) +
  geom_histogram(alpha = 0.4) +
  theme_classic() +
  xlab("Date of Service Provided") +
  ylab("Count") +
  ggtitle("Date when services were provided")

```


UMD was estabblished in 1983, so it is possible that data provided before this date might be not relevant or mistaken.
We can also see that almost *no information* provided for the dates before year 2000.

# Filter the data
```{r message=FALSE, warning=FALSE}

UMD_cld <- UMD_data %>%
  filter(between(N_Date, as.Date("2000-01-01"), as.Date("2019-06-30"))) %>%
  mutate(Year = as.numeric(format(N_Date, "%Y"))) 

```

# Total clients supported by year
```{r message=FALSE, warning=FALSE}

clients = UMD_cld %>%
  group_by('Client file number', Year) %>%
  summarize(N_Client=n())

# Print the result

ggplot(clients, aes(Year, N_Client)) + 
  geom_col(alpha = 0.4) +
  geom_smooth(color='red')+
  theme_classic()+
  xlab("Year of Service Provided") +
  ylab("Number of Clients") +
  ggtitle("Increase in number of clients by year")

```

# Food pounds provided per year
```{r message=FALSE, warning=FALSE}

food_df <- UMD_cld %>%
  drop_na(`Food Pounds`) %>%
  group_by(Year) %>%
  summarise(count = n(),food = sum(`Food Pounds`))

ggplot(food_df,aes(x = Year, y = food)) +
  geom_col(alpha = 0.4) +
  geom_smooth(color='red', se=FALSE) +
  theme_classic() +
  labs(title = 'Food (pounds) per Year',
       y = 'Food Pounds')

```

# How much food (in pounds) will be provided for next 5 years?
```{r message=FALSE, warning=FALSE}
food_model=lm(food ~ Year, food_df)
pred_years <- c(2019, 2020, 2021, 2022, 2023)

food_pred <- data.frame(Year=pred_years)
prd=predict(food_model, food_pred)

names(prd)[] <- c('2019','2020', '2021', '2022', '2023')

prd

```

# Distribution of diapers by year
```{r message=FALSE, warning=FALSE}
UMD_cld %>%
  drop_na('Diapers') %>%
  group_by(Year) %>%
  summarise(count = n(),diapers=sum(Diapers)) %>%
  ggplot(aes(x = Year, y = diapers)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_smooth(color='red', se=FALSE) +
  theme_classic() +
  labs(title = 'Diapers help by Year',
       y = 'Number of diapers')
```


# Distribution of clothing items by year

```{r message=FALSE, warning=FALSE}
UMD_cld %>%
  drop_na(`Clothing Items`) %>%
  group_by(Year) %>%
  summarise(count = n(),cloth=sum(`Clothing Items`)) %>%
  ggplot(aes(x = Year, y = cloth)) +
  geom_point(size = 1.5, alpha = 0.4) +
  geom_smooth(color='red', se=FALSE) +
  theme_classic() +
  labs(title = 'Clothing Items distribution by Year',
       y = 'Number of Clothing Items')

```


# Distribution of School kits by year
```{r message=FALSE, warning=FALSE}
UMD_cld %>%
  drop_na(`School Kits`) %>%
  group_by(Year) %>%
  summarise(count = n(), school=sum(`School Kits`)) %>%
  ggplot(aes(x = Year, y = school)) +
  geom_col(alpha = 0.4) +
  theme_classic() +
  labs(title = 'School Kits distribution by Year',
       y = 'Number of School Kits')

```


# Distribution of Hygiene kits by year

```{r message=FALSE, warning=FALSE}
UMD_cld %>%
  drop_na(`Hygiene Kits`) %>%
  group_by(Year) %>%
  summarise(count = n(), kit = sum(`Hygiene Kits`)) %>%
  ggplot(aes(x = Year, y = kit)) +
  geom_col(alpha = 0.4) +
  theme_classic() +
  labs(title = 'Hygiene Kits distribution by Year',
       y = 'Number of Hygiene Kits')

```


# Average family size (by column food provided for)

```{r message=FALSE, warning=FALSE}
family_size <- UMD_cld %>%
  select(`Food Provided for`) %>%
  filter(`Food Provided for`>0) %>%
  drop_na(`Food Provided for`) %>%
  rename(food=`Food Provided for`)

mean(family_size$food)

```

